---
title: "BC tools"
author: "Aleš Saska <mail@alessaska.cz>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# bctools package

This package contains pretty printer for R language Bytecode. It can be used for debugging purposes and for code analysis. 
It defines print method for bytecode object ( which is used internally by R language ).

Print function support 3 levels of verbosity ( parameter verbose = 0 or 1 or 2 ).

* verbose=0 - Print only source references ( if they are available, if they aren't print expression references instead )
* verbose=1 - The same as 0 + print bytecode version and print expression references ( if they are available ) and also print bytecode version
* verbose=2 - The same as 1 + print all operands' arguments ( including ones used just for debugging )

## Output description

* labels starts with 1:, 2: ... ( references to them begins with $ - $1 / $2 ... ).
* expression references begins with @
* source references with " - " ( if you want to be sure there are source references in bytecode included, call options(keep.source = TRUE) )

## Examples

In this short example you can see how to use the print method for bytecode object. You can see the bytecode object is 
created with the \emph{compiler} package ( which is built in the default R distribution ).
To see how create bytecode object visit [compiler manual page](https://stat.ethz.ch/R-manual/R-devel/library/compiler/html/compile.html).

```{r}

library(compiler)
library(bctools)

x <- 2

bytecode <- compiler::compile(quote(x+2));

print(compiler::disassemble(bytecode));

```

You can see that there is print method that you can call on bytecode object which is returned by compiler::disassemble function.


## Examples that show different verbosity levels

### Verbose 0 ( default )

In the following example you can see printed source references ( which begins with @ )

```{r}

library(compiler)
library(bctools)

f <- function(x){
    sum <- 0
    while(x > 0){
        sum <- sum + x
    }
    sum
}

bytecode <- compiler::cmpfun(f);

print(compiler::disassemble(bytecode));

```

### Verbose 1

In the following example you can see also printed expression references ( which begins with " - FILENAME:LINENO#" )

```{r}

library(compiler)
library(bctools)

f <- function(x){
    sum <- 0
    while(x > 0){
        sum <- sum + x
    }
    sum
}

bytecode <- compiler::cmpfun(f);

print(compiler::disassemble(bytecode), verbose=1);

```


### Verbose 2

The last example is the same code except with the verbosity level setted to 2 ( the highest possible ).
Notice that there occures for example aditional argument in ADD instruction ( it is used just for debugging purposes ).


This argument was hidden in lower levels of verbosity, because it is not normally used for the code evaluation.

```{r}

library(compiler)
library(bctools)

f <- function(x){
    sum <- 0
    while(x > 0){
        sum <- sum + x
    }
    sum
}

bytecode <- compiler::cmpfun(f);

print(compiler::disassemble(bytecode), verbose=2);


```
